---
title: "Introducing gtrendsR"
layout: post
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`gtrendsR` provides an interface for retrieving and displaying the information returned online by Google Trends. Trends (number of hits) over the time as well as geographic representation of the results can be displayed.

The package can be installed using the following command:

```{r, eval = FALSE}
devtools::install_github("PMassicotte/gtrendsR")
```

There are two main functions in the package: `gconnect()` to connect to Google using your credentials and `gtrends()` to perform actual queries.

```{r}
library(gtrendsR)
ls("package:gtrendsR")
```

# Examples

In this simple example, Googling trends for keywords `nhl`, `nba` and `nfl` are retrieved.

## Get the data

```{r, eval = FALSE}
library(gtrendsR)

usr <- "user@gmail.com"
psw <- "password"

gconnect(usr, psw)

sport_trend <- gtrends(c("nhl", "nba", "nfl"))
```

## Plot the data

```{r}
## Load sample data
data("sport_trend")

plot(sport_trend)

```

It is also possible to plot geographical data using `googleVis` as follow.

```{r, eval = FALSE}
plot(sport_trend, type = "region")
plot(sport_trend, type = "cities")
```

Feel free to try it!
