---
title: "R 101"
layout: post
tags:
  - R
  - R 101
---

Sometimes when you open a data file (lets say a *.csv*), variables will be recognized as *factor* whereas it should be numeric. It is therefore tempting to simply convert the variable to numeric using *as.numeric()*. Big mistake! If you use directly *as.numeric()* on a variable of the type *factor*, R will returns the levels of the factor rather the actual values. To overcome this hassle, you should first convert the variable into a string with *as.character()* and thereafter use *as.numeric()*.

```{r}
x <- as.factor(rnorm(5))
x

## Wrong way
as.numeric(x)

## Right way
as.numeric(as.character(x))

```

Another way to prevent numeric to be converted into factor would be to use *as.is = T* when opening the data file with *read.csv*.

## Remove all objects in the workspace**

```{r}
rm(list = ls())
```

## Replacing values of a data frame with NA


```{r, results = "asis", warning = FALSE}

library(xtable)

## Random matrix with with some (50) zeros
df <- matrix(rnorm(100), ncol = 10)

df[sample(100, 50)] <- 0

df <- data.frame(df)

print(xtable(df), type = "html")

## Replace the 0 with NA
df[df == 0] <- NA

print(xtable(df), type = "html", NA.string = "NA")
```

##Count % of missing values per column

The other day I was working with a large database. I was looking for a way to calculate % of missing values for each variable. Here is how I calculated it.

```{r, results = "asis"}

## Generate some data
df <- matrix(rnorm(100), ncol = 10)

df[sample(100, 50)] <- NA

df <- data.frame(df)

print(xtable(df), type = "html", NA.string = "NA")

## Find % of missing values in each column with apply
missing <- apply(df, 2, function(x) sum(is.na(x)))/nrow(df)*100

## Print the result
df2 <- data.frame(variable = names(df), missing = missing)

print(xtable(df2), type = "html", include.rownames = FALSE)
```

## Grouping data frame

Tsshere is various way to merge data frame in R. For me, the most easiest way is to use the join functions from the `dplyr` package.


<img src="{{ site.baseurl }}/assets/join_types.svg" />


```{r, results = "asis"}

library(dplyr)

df1 <- data.frame(ID = sample(LETTERS[1:20],10, replace = FALSE), x = rnorm(10))
df2 <- data.frame(ID = sample(LETTERS[1:20],10, replace = FALSE), y = rnorm(10))

print(xtable(df1), type = "html", include.rownames = FALSE)
print(xtable(df2), type = "html", include.rownames = FALSE)

inner_join(df1, df2, by = "ID")

df_exact_match <- merge(df1,df2, by = "ID") ## Exact match
print(xtable(df_exact_match), type = "html", include.rownames = FALSE, NA.string = "NA")

df_all_x <- merge(df1,df2, by = "ID", all.x = T) ## Keep all x (discard y that are not matched)
print(xtable(df_all_x), type = "html", include.rownames = FALSE, NA.string = "NA")

df_all_y <- merge(df1,df2, by = "ID", all.y = T) ## Keep all y (discard x that are not matched)
print(xtable(df_all_y), type = "html", include.rownames = FALSE, NA.string = "NA")

df_all <- merge(df1,df2, by = "ID", all = T) ## Keep everything
print(xtable(df_all), type = "html", include.rownames = FALSE, NA.string = "NA")
```
